---
title: "Stat 341 Homework 2"
author: "Name1 and Name2 and Name3"
date: '2019-03-04'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message=FALSE,warning=FALSE)
```

In this assignment we work with genotype data on 11 genetic markers
measured on 174 subjects. You do not need to know anything about
genotypes other than the fact that they take values 0, 1 or 2. 
The data are contained in a matrix with rows for subjects
and columns for markers in the file `snpMatrix.csv`.
The first column of `snpMatrix.csv` is the name of 
the subject and the other 11 columns are the marker
genotype data.
Further information on the markers is contained in `snpSupport.csv` and further
information on the subjects is contained in `subjectSupport.csv`.

Read in the data with the following code chunk.

```{r}
snpMatrix <- read.csv("snpMatrix.csv",stringsAsFactors = FALSE)
snpSupport <- read.csv("snpSupport.csv", stringsAsFactors = FALSE)
subjectSupport <- read.csv("subjectSupport.csv", stringsAsFactors = FALSE)
```

1. (2 marks) Create a "long" version of the `snpMatrix` data
called `snpLong`
with a key variable named `snp` and a value variable named `genotype` 
comprised of the marker genotypes.
```{r}
library(tidyverse)
snpLong <- gather(snpMatrix, snp, genotype, rs28412942:rs28358572, na.rm=TRUE)
head(snpLong)
```

2. (2 marks) Taken together, `snpLong`, `snpSupport` 
and `subjectSupport` are a relational dataset. What are
the names of the primary keys of the `snpSupport` and `subjectSupport`
data frames?

_Answer_
snp for snpSupport, Sample.name for subjectSupport.

3. (2 marks) Create a SQLite database called `snp` that contains
the `snpLong`, `snpSupport` 
and `subjectSupport` dataframes as its tables. List the tables
and then remove the database connection and the file on
your computer.
(Note: We will **not** use this database in the questions to follow.)
```{r}
library(DBI)
library(RSQLite)
snp <- dbConnect(RSQLite::SQLite(), 'snp.sqlite') 
dbWriteTable(snp, name = "snpLong", value = snpLong, overwrite = TRUE)
dbWriteTable(snp, name = "snpSupport", value = snpSupport, overwrite = TRUE)
dbWriteTable(snp, name = "subjectSupport", value = subjectSupport, overwrite = TRUE)
dbListTables(snp)
dbDisconnect(snp)
unlink('snp.sqlite')
```

4. (3 marks)
Starting from the dataframes `snpLong`, `snpSupport` and
`subjectSupport`, use functions from dplyr and stringr
to produce a dataframe with columns `Sample.name`, `snp`
and `genotype` for (i) subjects listed as `female` in `subjectSupport`
and (ii) markers that have the string `SNP` in the `assayLSID` column
of `snpSupport`. Save your results in a dataframe called
`snpFemaleSNP` and print out the first six rows
of `snpFemaleSNP` with the `head()` function.

```{r}
library(stringr)
subjectSupport %>% filter(Sex == "female") -> i
pattern <- "SNP"
snpSupport %>% filter(str_detect(snpSupport$assayLSID, 'SNP')) -> ii
join <- inner_join(snpLong, i, by = 'Sample.name')
join2 <- inner_join(join, ii, by = 'snp')
join2 %>% select(Sample.name, snp, genotype) -> snpFemaleSNP
head(snpFemaleSNP)
```



5. (0 marks) We did not discuss the `spread()` function in class.
It is the opposite of `gather()`. Cut-and-paste
the following code chunk to produce a dataset with rows for 
subjects and columns for markers based on `snpFemaleSNP`

```{r}
dat <- spread(snpFemaleSNP,key=snp,value=genotype)
head(dat)
```