---
title: "Homework 3"
author: "Name1, Name2, Name3"
date: '2019-03-25'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages

```{r, message=FALSE, warning=FALSE}
library(ggplot2)
library(dplyr)
library(stringr)
library(nycflights13)
library(DBI)
```

## Questions

0. (0 marks) If you haven't already, install
the `nycflights13` package and read the 
help files on the 
`flights` and `airlines` tables.
```{r}
data(flights)
data(airlines)
head(flights)
head(airlines)
```
1. (1 mark) 
Use `dplyr` functions to perform an inner join between the `airlines` table
and the `flights` table on `carrier` and 
then select the variables
`carrier`, `name`, `month`, `day`, `dep_time`
and `dep_delay`.
Save your dataframe in `flights`.

```{r}
flights %>% inner_join(airlines, by = 'carrier') %>% select(carrier, name, month, day, dep_time, dep_delay) -> flights
head(flights)
```

2. (2 marks) Remove Inc., Co. and Corporation from the airline names in the
`name` variable of `flights`. Then shorten the phrases "Airlines",
"Air Lines", and "Airways" to just "Air". Print out the 
unique values of `flights$name`.

```{r}
pattern1 <- "Inc."
pattern2 <- 'Co.'
pattern3 <- "Corporation"
pattern4 <- 'Airlines'
pattern5 <- "Air Lines"
pattern6 <- "Airways"
flights$name %>% str_replace_all(pattern = pattern1, replacement = "") %>% str_replace_all(pattern2, "") %>% str_replace_all(pattern3, "") %>% str_replace_all(pattern4, "Air") %>% str_replace_all(pattern5, "Air") %>% str_replace_all(pattern6, "Air") -> flights$name
unique(flights$name)
```

3. (2 marks) Select the flights from January. Calculate the median departure 
delay for each airline on each day and save the results in a 
dataframe called fldelaysJan. Plot median delay versus day, 
with different facets for each airline. (Note that
the `median()` function in R has an argument `na.rm`
that you may wish to set to `FALSE`.) 
Use the `name` variable
to specify airlines (instead of `carrier`) so that the labels
on your facets are informative. Do **not** free
the scales on the facets.

```{r, message=FALSE}
# Make data and lists for function inputs
library(tidyverse)
flights %>% filter(month == 1) %>% select(name, dep_delay, day)-> Jan_flt
names <- as.list(unique(Jan_flt$name))
days <- as.list(1:31)

# Make a function to calculate median delays by specifying  variable 'day'
med_calculator <- function(d) {
    fs <- vector("double", length(names))
    for (i in seq_along(names)) {
      Jan_flt %>% filter(day == d) %>% filter(name == names[[i]]) %>% select(dep_delay) %>% unlist() %>% median(na.rm = TRUE) -> fs[i]
  }
  fs
}

# Loop through all days in January and stack results
fldelaysJan <- data.frame()
for (j in seq_along(days)){
   df <- data.frame(day = j, name = as.vector(unique(Jan_flt$name)), med_del = med_calculator(j), stringsAsFactors = FALSE)
   fldelaysJan <- rbind(fldelaysJan, df)
}
head(fldelaysJan)
```
```{r}
#Plot
ggplot(fldelaysJan, aes(x = day, y = med_del, color = name )) + geom_line() + facet_wrap(~name)
```

4. (2 marks) The plot in part (3) is not very informative because one airline 
had a very long median flight delay on one day in January. Remove this 
**one observation** from `fldelaysJan`. Also remove 
the airline
that had flights on only one day. Now redo the plot in (3).
Again, do **not** free the scales on the facets.

```{r}
outlier <- max(fldelaysJan$med_del, na.rm = TRUE) # Find the one outlier
fldelaysJan <- fldelaysJan[-which(fldelaysJan$med_del == outlier), ]
fldelaysJan <- fldelaysJan[-which(fldelaysJan$name == "SkyWest Air "), ]
ggplot(fldelaysJan, aes(x = day , y =med_del, color = name)) + geom_line() + facet_wrap(~name)
```
